An Analysis of Security Vulnerabilities in Container Images for Scientific Data Analysis

### 2020-10-26 paper.tex
edits in abstract and intro

\begin{abstract}
Software containers greatly facilitate the deployment and reproducibility
of scientific data analyses on high-performance computing clusters
(HPC). However, container images often contain outdated or unnecessary
software packages, which increases the number of security vulnerabilities
in the images and widens the attack surface of the infrastructure. This
paper presents a vulnerability analysis of container images for scientific
data analysis. We compare results obtained with four vulnerability
scanners, focusing on the use case of neuroscience data analysis, and
quantifying the effect of images updates and minification on the number of
vulnerabilities. We find that container images used for neuroscience data analysis
contain hundreds of vulnerabilities, that software updates remove about two
thirds of these vulnerabilities, and that removing unused packages is also
effective. We conclude with recommendations on how to build container
images with a reduced amount of vulnerabilities.

@@

Container images typically include full operating system (OS) distributions in
addition to data analysis software and their dependencies. They are rarely
updated due to concerns that software updates will interfere with
the results~\cite{gronenschild2012effects, glatard2015reproducibility}.
Images also typically include more dependencies than required, to make them
easier to reuse between experiments. As a result, over 30\% of official
images in DockerHub have been shown to contain high priority security
vulnerabilities~\cite{gummaraju2015over}, images on average contain over
180 vulnerabilities~\cite{Shu2017}, and vulnerabilities are often caused by
outdated packages~\cite{zerouali2019relation}.

In this study, we focus on the vulnerabilities present in container images
deployed on HPC clusters for scientific data analysis. We address the
following questions:

@@

\textit{Can the amount of vulnerabilities be reduced by updating the images?}  
For many reasons related to reproducibility and lifecycle of research
projects, container images deployed on HPC clusters often include outdated
software. We report on the effect of software updates on the amount of
vulnerabilities found in images.

\textit{Can the amount of vulnerabilities be reduced by minifying images?} 
Container images often include more software packages than necessary for 
typical analysis. We report on the impact of unused software packages on
the presence of vulnerabilities.

The remainder of this paper 
describes the materials and tools used in the
experiments, including container images, image scanners, and our methodology for updating and minifying images.
Results present the vulnerabilities detected in container
images, quantify the effectiveness of updating and minifying images, and
explain the differences observed between scanners. In conclusion, we
provide a set of image creation guidelines for a more secure deployment of
containers on HPC clusters.

>>>>

\begin{abstract}
Software containers greatly facilitate the deployment and reproducibility
of scientific data analyses on high-performance computing clusters
(HPC). However, container images often contain outdated or unnecessary
software packages, which increases the number of security vulnerabilities
in the images and widens the attack surface of the infrastructure. This
paper presents a vulnerability analysis of container images for scientific
data analysis. We compare results obtained with four vulnerability
scanners, focusing on the use case of neuroscience data analysis, and
quantifying the effect of image update and minification on the number of
vulnerabilities. We find that container images used for neuroscience data analysis
contain hundreds of vulnerabilities, that software updates remove about two
thirds of these vulnerabilities, and that removing unused packages is also
effective. We conclude with recommendations on how to build container
images with a reduced amount of vulnerabilities.

@@

Container images typically include full operating system (OS) distributions in
addition to data analysis software and their dependencies. They are rarely
updated due to concerns that software updates will interfere with
the results~\cite{gronenschild2012effects, glatard2015reproducibility}.
Images also typically include more dependencies than required, to make them
easier to reuse between experiments. As a result, over 30\% of official
images in DockerHub have been shown to contain high-priority security
vulnerabilities~\cite{gummaraju2015over}, images on average contain over
180 vulnerabilities~\cite{Shu2017}, and vulnerabilities are often caused by
outdated packages~\cite{zerouali2019relation}.

In this study, we focus on the vulnerabilities present in container images
deployed on HPC clusters for scientific data analysis, in particular in the
neuroimaging domain. We address the following questions:

@@

\textit{Can the amount of vulnerabilities be reduced by updating the images?}  
For many reasons related to reproducibility and the lifecycle of research
projects, container images deployed on HPC clusters often include outdated
software. We report on the effect of software updates on the amount of
vulnerabilities found in images.

\textit{Can the amount of vulnerabilities be reduced by minifying images?} 
Container images often include more software packages than necessary for 
a typical analysis. We report on the impact of unused software packages on
the presence of vulnerabilities.

The remainder of this paper 
describes the container images and scanners used in our experiment, and our methodology for updating and minifying images.
Results present the vulnerabilities detected in container
images, quantify the effectiveness of updating and minifying images, and
explain the differences observed between scanners. In conclusion, we
provide a set of image creation guidelines for a more secure deployment of
containers on HPC clusters.
